---
title: "GSS challange More Data"
author: "Daryl P"
date: "FALL 2022"
output: html_document
---







```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS2016.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,cappun)
mydata <- na.omit(mydata)
```

# Introduction

```{r eval=FALSE}
mydata<- mydata %>%
mutate(across(where(is.character), as.factor))
```
<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is the attitude toward capitol punishment (cappun) independent of race (race)?
If not, then whose view differs from whom?

To establish the answer to this statement requires the use of numerical, graphical and inferential data mentioned in the methods below to support our decisions.

# Methods
The types of variables in this problem consist of races and capital punishment both of which are catagorical variable files. therefoe cat cat helper files will be a good choice to assist in analyizing the Race and capital  punishment data.

<!--Decide on your methods:  use "variable aonalysis" or other appropriate
descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical  and inferential tools will be used!-->
the variable analysis is best show by using the chart plot that shows the frequency and percentages of the people used in the collection of data.

# Results are shown thru the use of numerical and graphical data that more whites are for capital punishment than blacks or other races

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results
looking at the plot it appears their is a difference in how different races show their choice about(cappun)

### Graphical Descriptive Results
The barchart below is a graphical representation of the the people separated by race and there choices, those in favor or oppsed to capital punishment

``` is a graphical representation
barchartGC(~race + cappun,data=mydata)
barchartGC(~race + cappun,data=mydata, type="percent

graphical chart descriptionT


## Now with More data



<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
table1 <- xtabs(~race + cappun, data=mydata)
rowPerc(table1)
colPerc(table1)
```


The numerical results display that 65% of whites favor capital punishment and  55% of other races are in favor of capital punishment. and 58% blacks are oppsed.



## Inferential Results
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
THE DATA BELOW from the fisher XY and 
```{r}
fisherXY(mydata$race,mydata$cappun)
```
`


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The P value is 0. This indicates that there is a statistical difference when  referring to people and there choices toward capital punishment. the Fisher XY test is needed to  consider multible variables the chi-square test is not enough,.since we have Three.


```{r}
fisherXY(mydata$race,mydata$cappun)
```



The fisher xy function test shows that there is a statistical difference on the choices that were  made by white and black races in regards to capital punishment are different according to the p value of the fisher xy test results, how ever when we got more data the p value shows more difference between each race and a bigger difference is between the whites and the blacks, because of a lower p.value.

# Discussion and Conclusion


## More data version



<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
